Evolution of Processes in Delivering Reliable Software within the European Scientific Arena

### 2018-09-27 paper-ieee-trans.tex
Review #2

@@

\begin{figure*}
	\centering
	\begin{subfigure}
		\centering
\includegraphics[width=0.85\textwidth]{images/disvis-flow.png}
\caption{DevOps pipeline to distribute Docker images for Disvis application.}
\label{fig:fig_disvis}
	\end{subfigure}
	\quad
	\begin{subfigure}
		\centering
\includegraphics[width=\textwidth]{images/pilotpreview.png}
\caption{Resource centers supporting the Pilot Preview testbed and corresponding
set of deployed {\sl INDIGO--DataCloud} components or services.}
\label{fig:fig_pilotpreview}
	\end{subfigure}
\end{figure*}

@@

% Comment
%%P8 C1 L26: What is the demonstration that reliability has been improved significantly with the addition of stricter methologies? Is there some quantitative evidence of uptime, or bug reports per user, or similar? Or some evidence of speed of defect detection & fixing, which you cite as greatly improved? I am very sympathetic to the idea that systematic development techniques can help improve software quality (without getting into circular definitions, where "quality" means that methodologies or standards have been applied!) but there will also be a point at which the methodology gets too heavy and compromises the "agility" of the development team: how can you identify which processes are helping and which are not?
Reliability of the software has been substantially improved by applying the SQA
criteria in the development stage. Consequently, each change in the source code
that is meant to be integrated in the production version, is validated according
to the SQA testing requirements. Thus, defects are detected and corrected early in the
software lifecycle, reducing the cost of fixing errors at later stages. Automation is
key in this process, allowing the implementation of a CI scenario that enforces
the SQA criteria compliance. The application of more advanced DevOps practices
resulted in the implementation of CD pipelines to extend automation to the
delivery of packages. Thereby, safe and frequent software releases are possible, reducing
the impact on production systems. In this regard, a pre--production validation,
through the preview testbeds or resource center's early adoption -- handled by the
Staged--rollout process -- alleviates the likelihood of discovering uncovered
issues at production. To put it in numbers, as Fig.~\ref{fig:ingigobugs} showcases, the 
total number of defects detected in the pre--production stages surpassed by a factor of 30
the ones reported by external users, over the lifetime of the {\sl INDIGO-DataCloud} 
project.

@@

\begin{figure}[ht]
\centering
\includegraphics[width=0.5\textwidth, height=50mm]{images/indigoBugs.png}
\caption{Number of defects detected in development and production stages over the lifetime of the
{\sl INDIGO-DataCloud} project. The defects started to arise in the second year of the project, 
in parallel with the release of {\sl INDIGO--1} software, while the major activity was concentrated in the
months that preceded the second major release of {\sl INDIGO--2}, in April 2017. The practical implementation
of the SQA definition described throughout this paper was the key actor in uncovering defects while in pre--production
stages, thus not affecting the software released.}
\label{fig:indigobugs}
\end{figure}

>>>>

\begin{figure*}[ht]
	\centering
	\begin{subfigure}
		\centering
\includegraphics[width=0.85\textwidth]{images/disvis-flow.png}
\caption{DevOps pipeline to distribute Docker images for Disvis application.}
\label{fig:fig_disvis}
	\end{subfigure}
	\quad
	\begin{subfigure}
		\centering
\includegraphics[width=\textwidth]{images/pilotpreview.png}
\caption{Resource centers supporting the Pilot Preview testbed and corresponding
set of deployed {\sl INDIGO--DataCloud} components or services.}
\label{fig:fig_pilotpreview}
	\end{subfigure}
\end{figure*}

@@

@@

% Comment
%%P8 C1 L26: What is the demonstration that reliability has been improved significantly with the addition of stricter methologies? Is there some quantitative evidence of uptime, or bug reports per user, or similar? Or some evidence of speed of defect detection & fixing, which you cite as greatly improved? I am very sympathetic to the idea that systematic development techniques can help improve software quality (without getting into circular definitions, where "quality" means that methodologies or standards have been applied!) but there will also be a point at which the methodology gets too heavy and compromises the "agility" of the development team: how can you identify which processes are helping and which are not?
Reliability of the software has been substantially improved by applying the SQA
criteria in the development stage. Consequently, each change in the source code
that is meant to be integrated in the production version, is validated according
to the SQA testing requirements. Thus, defects are detected and corrected early in the
software lifecycle, reducing the cost of fixing errors at later stages. Automation is
key in this process, allowing the implementation of a CI scenario that enforces
the SQA criteria compliance. The application of more advanced DevOps practices
resulted in the implementation of CD pipelines to extend automation to the
delivery of packages. Thereby, safe and frequent software releases are possible, reducing
the impact on production systems. In this regard, a pre--production validation,
through the preview testbeds or resource center's early adoption -- handled by the
Staged--rollout process -- alleviates the likelihood of discovering uncovered
issues at production. To put it in numbers, as Fig.~\ref{fig:indigobugs} showcases, the
total number of defects detected in the pre--production stages surpassed by a factor of 30
the ones reported by external users, over the lifetime of the {\sl INDIGO-DataCloud}
project.
\begin{figure}[h]
\centering
\includegraphics[width=0.5\textwidth, height=50mm]{images/indigoBugs.png}
\caption{Number of defects detected in development and production stages over the lifetime of the
{\sl INDIGO-DataCloud} project. The defects started to arise in the second year of the project,
in parallel with the release of {\sl INDIGO--1} software, while the major activity was concentrated in the
months that preceded the second major release of {\sl INDIGO--2}, in April 2017. The practical implementation
of the SQA definition described throughout this paper was the key actor in uncovering defects while in pre--production
stages, thus not affecting the software released.}
\label{fig:indigobugs}
\end{figure}

@@
